Inference for trans-dimensional Bayesian models with Diffusive Nested Sampling

### 2014-06-27 Documents/Paper/ms.tex
Nested sampling etc

Here we have assumed the priors for $N$ and $\hyperparams$ are independent, and
the interim prior for $\{\xx_i\}$ is iid and does not depend on $N$.

@@

In other studies, it has been common to do separate runs of Nested Sampling
with different values of $N$. Then, the posterior for $N$ can be calculated
based on the estimates of the evidence or marginal likelihood. Our motivation
for using Nested Sampling is different. With reversible jump MCMC it is possible
to obtain the posterior for $N$ with a single run. However, the exploration
of the posterior may be very difficult, and sometimes the problem may even
contain a phase transition. We use Nested Sampling to overcome these
difficulties.

>>>>

Here we have assumed the priors for $N$ and $\hyperparams$ are independent, and
the interim prior for $\{\xx_i\}$ is iid and does not depend on $N$.
Given data $\mathcal{D}$, we usually want to generate samples from the
posterior distribution given by:
\begin{eqnarray}
p(N, \hyperparams, \{\xx_i\} | \mathcal{D}) \propto
p(N, \hyperparams, \{\xx_i\})
p(\mathcal{D} | N, \hyperparams, \{\xx_i\})\label{eq:bayes}
\end{eqnarray}
The marginal likelihood, or evidence, is also an important quantity, and is
the normalisation constant of the posterior given in Equation~\ref{eq:bayes}:
\begin{eqnarray}
\mathcal{Z} &=& p(\mathcal{D})\\
&=& \sum_{i=0}^{N_{\rm max}} \int
p(N, \hyperparams, \{\xx_i\})p(\mathcal{D} | N, \hyperparams, \{\xx_i\})
\, d^N \xx_i \, d\hyperparams \label{eq:evidence}
\end{eqnarray}

@@

In other studies, it has been common to do separate runs of Nested Sampling
$N$ fixed at various trial values. Then, the posterior for $N$ can be calculated
based on the estimates of the evidence obtained as a function of $N$. However,
in this paper, we use trans-dimensional MCMC moves to infer $N$. In other words,
$N$ is treated as just another parameter, and the marginal likelihood defined
in Equation~\ref{eq:evidence}, which we can compute, also includes the sum over
the unknown parameter $N$. Our motivation
for using Nested Sampling is not primarily to calculate $\mathcal{Z}$.
The motivation for using Nested Sampling, rather than just sampling the
posterior using trans-dimensional MCMC, is that the posterior often contains
difficult features, such as strong dependencies or multiple modes, that cause
problems with mixing. Diffusive Nested Sampling
(DNS) replaces the posterior distribution with an alternative target
distribution composed of a mixture of the prior distribution with other, more
constrained distributions. This is similar to the idea of simulated annealing,
where the target distribution is modified to something ``easier'', however,
the sequence of distributions used in Nested Sampling avoids some of the
problems with simulated annealing: in particular, there is no need to choose
a temperature schedule, and the method does not fail when phase transitions
are present.
